---
title: "Blog"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
```


```{r load packs}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, DT, stringr, purrr, dplyr, ggplot2, plotly)
```


```{r getting stats, echo=FALSE}
Season2018 <- readRDS("Season2018.rds")
TeamStats2018 <- Season2018 %>% mutate(EVENT_TYPE = make.names(EVENT_TYPE), SHOT_TYPE = str_replace(SHOT_TYPE, " Field Goal", ""))  %>% group_by(TEAM_NAME, SHOT_TYPE, EVENT_TYPE) %>% dplyr::summarise(N = n()) %>% spread(key = EVENT_TYPE, value = N) %>% split(.$SHOT_TYPE) %>%  map(~mutate(.x, ShotPct = round(Made.Shot/(Made.Shot +Missed.Shot), 3), Total = Made.Shot + Missed.Shot))

Points <- c("2Pts", "3Pts")

for(i in 1:length(TeamStats2018)){
  colnames(TeamStats2018[[i]])[2:6] <- paste0(colnames(TeamStats2018[[i]])[2:6], Points[i])
}

TeamStats2018 <- TeamStats2018 %>% reduce(merge) %>% select(-SHOT_TYPE2Pts, -SHOT_TYPE3Pts, -Made.Shot2Pts, -Made.Shot3Pts, -Missed.Shot2Pts, -Missed.Shot3Pts) %>% mutate(TotalShots = Total2Pts + Total3Pts) %>% mutate(PercentageOf2s = round(Total2Pts/TotalShots,2),PercentageOf3s = round(Total3Pts/TotalShots, 2)) %>% select(-TotalShots, -Total2Pts, -Total3Pts)
Teams <- unique(Season2018$TEAM_NAME)

Prop <- data.frame(TEAM_NAME = Teams , PropShot = NA)

for(i in 1:30){
Offa <- dplyr::filter(Season2018, HTM == Teams[i] | VTM == Teams[i])
Prop$PropShot[i] <- round(nrow(dplyr::filter(Offa, TEAM_NAME == Teams[i]))/nrow(dplyr::filter(Offa,TEAM_NAME != Teams[i])),3)
}

TeamStats2018 <- full_join(TeamStats2018, Prop) %>% mutate(PPS = round((ShotPct2Pts*2*PercentageOf2s)+(ShotPct3Pts*3*PercentageOf3s),3)) %>% mutate(AdjPPS = round((PPS * PropShot),3)) %>% mutate(Season = 2018)
```

```{r MakeTable, echo=FALSE}
TeamStats2018 <- TeamStats2018
```

```{r, cache = TRUE}
pacman::p_load(tidyverse, DT, stringr, purrr, dplyr)
library(SpatialBall)

Season2001 <- readRDS("Season2001.rds")
TeamStats2001 <- Season2001 %>% mutate(EVENT_TYPE = make.names(EVENT_TYPE), SHOT_TYPE = str_replace(SHOT_TYPE, " Field Goal", ""))  %>% group_by(TEAM_NAME, SHOT_TYPE, EVENT_TYPE) %>% dplyr::summarise(N = n()) %>% spread(key = EVENT_TYPE, value = N) %>% split(.$SHOT_TYPE) %>%  map(~mutate(.x, ShotPct = round(Made.Shot/(Made.Shot +Missed.Shot), 3), Total = Made.Shot + Missed.Shot))

Points <- c("2Pts", "3Pts")

for(i in 1:length(TeamStats2001)){
  colnames(TeamStats2001[[i]])[2:6] <- paste0(colnames(TeamStats2001[[i]])[2:6], Points[i])
}

TeamStats2001 <- TeamStats2001 %>% reduce(merge) %>% select(-SHOT_TYPE2Pts, -SHOT_TYPE3Pts, -Made.Shot2Pts, -Made.Shot3Pts, -Missed.Shot2Pts, -Missed.Shot3Pts) %>% mutate(TotalShots = Total2Pts + Total3Pts) %>% mutate(PercentageOf2s = round(Total2Pts/TotalShots,2),PercentageOf3s = round(Total3Pts/TotalShots, 2)) %>% select(-TotalShots, -Total2Pts, -Total3Pts)
Teams <- unique(Season2001$TEAM_NAME)

Prop <- data.frame(TEAM_NAME = Teams , PropShot = NA)

for(i in 1:length(Teams)){
  Offa <- dplyr::filter(Season2001, HTM == Teams[i] | VTM == Teams[i])
  Prop$PropShot[i] <- round(nrow(dplyr::filter(Offa, TEAM_NAME == Teams[i]))/nrow(dplyr::filter(Offa,TEAM_NAME != Teams[i])),3)
}

TeamStats2001 <- full_join(TeamStats2001, Prop) %>% mutate(PPS = round((ShotPct2Pts*2*PercentageOf2s)+(ShotPct3Pts*3*PercentageOf3s),3)) %>% mutate(AdjPPS = round((PPS * PropShot),3)) %>% mutate(Season = 2001)

TeamStats2001 <- TeamStats2001

data("season2017")
Season2017 <- season2017
TeamStats2017 <- Season2017 %>% mutate(EVENT_TYPE = make.names(EVENT_TYPE), SHOT_TYPE = str_replace(SHOT_TYPE, " Field Goal", ""))  %>% group_by(TEAM_NAME, SHOT_TYPE, EVENT_TYPE) %>% dplyr::summarise(N = n()) %>% spread(key = EVENT_TYPE, value = N) %>% split(.$SHOT_TYPE) %>%  map(~mutate(.x, ShotPct = round(Made.Shot/(Made.Shot +Missed.Shot), 3), Total = Made.Shot + Missed.Shot))

Points <- c("2Pts", "3Pts")

for(i in 1:length(TeamStats2017)){
  colnames(TeamStats2017[[i]])[2:6] <- paste0(colnames(TeamStats2017[[i]])[2:6], Points[i])
}

TeamStats2017 <- TeamStats2017 %>% reduce(merge) %>% select(-SHOT_TYPE2Pts, -SHOT_TYPE3Pts, -Made.Shot2Pts, -Made.Shot3Pts, -Missed.Shot2Pts, -Missed.Shot3Pts) %>% mutate(TotalShots = Total2Pts + Total3Pts) %>% mutate(PercentageOf2s = round(Total2Pts/TotalShots,2),PercentageOf3s = round(Total3Pts/TotalShots, 2)) %>% select(-TotalShots, -Total2Pts, -Total3Pts)
Teams <- unique(Season2017$TEAM_NAME)

Prop <- data.frame(TEAM_NAME = Teams , PropShot = NA)

for(i in 1:length(Teams)){
  Offa <- dplyr::filter(Season2017, HTM == Teams[i] | VTM == Teams[i])
  Prop$PropShot[i] <- round(nrow(dplyr::filter(Offa, TEAM_NAME == Teams[i]))/nrow(dplyr::filter(Offa,TEAM_NAME != Teams[i])),3)
}

TeamStats2017 <- full_join(TeamStats2017, Prop) %>% mutate(PPS = round((ShotPct2Pts*2*PercentageOf2s)+(ShotPct3Pts*3*PercentageOf3s),3)) %>% mutate(AdjPPS = round((PPS * PropShot),3)) %>% mutate(Season = 2017)

TeamStats2017 <- TeamStats2017


pacman::p_load(tidyverse, DT, stringr, purrr, dplyr)

Season2016 <- readRDS("shotDataTotal2016.rds")
TeamStats2016 <- Season2016 %>% mutate(EVENT_TYPE = make.names(EVENT_TYPE), SHOT_TYPE = str_replace(SHOT_TYPE, " Field Goal", ""))  %>% group_by(TEAM_NAME, SHOT_TYPE, EVENT_TYPE) %>% dplyr::summarise(N = n()) %>% spread(key = EVENT_TYPE, value = N) %>% split(.$SHOT_TYPE) %>%  map(~mutate(.x, ShotPct = round(Made.Shot/(Made.Shot +Missed.Shot), 3), Total = Made.Shot + Missed.Shot))

Points <- c("2Pts", "3Pts")

for(i in 1:length(TeamStats2016)){
  colnames(TeamStats2016[[i]])[2:6] <- paste0(colnames(TeamStats2016[[i]])[2:6], Points[i])
}

TeamStats2016 <- TeamStats2016 %>% reduce(merge) %>% select(-SHOT_TYPE2Pts, -SHOT_TYPE3Pts, -Made.Shot2Pts, -Made.Shot3Pts, -Missed.Shot2Pts, -Missed.Shot3Pts) %>% mutate(TotalShots = Total2Pts + Total3Pts) %>% mutate(PercentageOf2s = round(Total2Pts/TotalShots,2),PercentageOf3s = round(Total3Pts/TotalShots, 2)) %>% select(-TotalShots, -Total2Pts, -Total3Pts)
Teams <- unique(Season2016$TEAM_NAME)

Prop <- data.frame(TEAM_NAME = Teams , PropShot = NA)

for(i in 1:length(Teams)){
  Offa <- dplyr::filter(Season2016, HTM == Teams[i] | VTM == Teams[i])
  Prop$PropShot[i] <- round(nrow(dplyr::filter(Offa, TEAM_NAME == Teams[i]))/nrow(dplyr::filter(Offa,TEAM_NAME != Teams[i])),3)
}

TeamStats2016 <- full_join(TeamStats2016, Prop) %>% mutate(PPS = round((ShotPct2Pts*2*PercentageOf2s)+(ShotPct3Pts*3*PercentageOf3s),3)) %>% mutate(AdjPPS = round((PPS * PropShot),3))%>% mutate(Season = 2016)

TeamStats2016 <- TeamStats2016

pacman::p_load(tidyverse, DT, stringr, purrr, dplyr)

Season2015 <- readRDS("shotDataTotal2015.rds")
TeamStats2015 <- Season2015 %>% mutate(EVENT_TYPE = make.names(EVENT_TYPE), SHOT_TYPE = str_replace(SHOT_TYPE, " Field Goal", ""))  %>% group_by(TEAM_NAME, SHOT_TYPE, EVENT_TYPE) %>% dplyr::summarise(N = n()) %>% spread(key = EVENT_TYPE, value = N) %>% split(.$SHOT_TYPE) %>%  map(~mutate(.x, ShotPct = round(Made.Shot/(Made.Shot +Missed.Shot), 3), Total = Made.Shot + Missed.Shot))

Points <- c("2Pts", "3Pts")

for(i in 1:length(TeamStats2015)){
  colnames(TeamStats2015[[i]])[2:6] <- paste0(colnames(TeamStats2015[[i]])[2:6], Points[i])
}

TeamStats2015 <- TeamStats2015 %>% reduce(merge) %>% select(-SHOT_TYPE2Pts, -SHOT_TYPE3Pts, -Made.Shot2Pts, -Made.Shot3Pts, -Missed.Shot2Pts, -Missed.Shot3Pts) %>% mutate(TotalShots = Total2Pts + Total3Pts) %>% mutate(PercentageOf2s = round(Total2Pts/TotalShots,2),PercentageOf3s = round(Total3Pts/TotalShots, 2)) %>% select(-TotalShots, -Total2Pts, -Total3Pts)
Teams <- unique(Season2015$TEAM_NAME)

Prop <- data.frame(TEAM_NAME = Teams , PropShot = NA)

for(i in 1:length(Teams)){
  Offa <- dplyr::filter(Season2015, HTM == Teams[i] | VTM == Teams[i])
  Prop$PropShot[i] <- round(nrow(dplyr::filter(Offa, TEAM_NAME == Teams[i]))/nrow(dplyr::filter(Offa,TEAM_NAME != Teams[i])),3)
}

TeamStats2015 <- full_join(TeamStats2015, Prop) %>% mutate(PPS = round((ShotPct2Pts*2*PercentageOf2s)+(ShotPct3Pts*3*PercentageOf3s),3)) %>% mutate(AdjPPS = round((PPS * PropShot),3))%>% mutate(Season = 2015)

TeamStats2015 <- TeamStats2015


```



```{r}
library(plotly)
TeamStats <- rbind(TeamStats2001, TeamStats2018)
TeamStats <- rbind(TeamStats, TeamStats2017)
TeamStats <- rbind(TeamStats, TeamStats2016)
TeamStats <- rbind(TeamStats, TeamStats2015)
TeamStats <- arrange(TeamStats, desc(PercentageOf3s))


p <- ggplot(TeamStats, aes(x = PercentageOf3s, y = PPS)) + geom_point() + geom_smooth(method = "lm") + ggtitle("Figure 1") + geom_text(aes(label=TEAM_NAME),hjust=0, vjust=0) + xlab("Proportion of shots that are 3pts")
 
```

## The first team ever to shoot more than half their shots beyond the three point line

For while we`ve been wondering if there is such a thing as too many three pointers for a team. Well we can keep asking ourselves that, as today, December the second of 2017, the Houston Rockets are the first team to ever shoot over half their shots from beyond the three point line. And it's not even close.

In the plot and table bellow we can see that this year the Rockets are shooting `r TeamStats$PercentageOf3s[1]*100` % of their shots from the three point range, this year, followed by the same team of last year. The third place, is last year's Cavs team, with `r TeamStats$PercentageOf3s[3]*100` %. That is a `r TeamStats$PercentageOf3s[1]*100 -TeamStats$PercentageOf3s[3]*100`% difference!!

### Why keep jacking up threes?

Does this make any sense? As we see in figure 1, there is a positive relationship between the percentage of shots a team takes that are three pointers with their points per shot. Actually, if we don't count the Golden State Warriors teams from 2016, 2017 and this year incarnation of the team, this year Rockets team has the highest Point per shot mark in the last 5 years at `r TeamStats$PPS[1]`, while making only `r TeamStats$ShotPct3Pts[1]*100`% of them.


```{r}
ggplotly(p)
```

### What would happend if the warriors shot as many threes as the Rockets

As told above, this year warriors are the top team in the last 5 years in Points per shot making `r dplyr::filter(TeamStats, TEAM_NAME == "GSW" & Season == 2018)$PPS`. Points per shot. They are making an efficient `r dplyr::filter(TeamStats, TEAM_NAME == "GSW" & Season == 2018)$ShotPct3Pts*100` of their three pointers and an outstanding `r dplyr::filter(TeamStats, TEAM_NAME == "GSW" & Season == 2018)$ShotPct2Pts*100` o their two pointers. But what would happen if they shot they same proportion of their shots form the three point line as the Rockets did?

Well, they would change their PPS to `r round((dplyr::filter(TeamStats, TEAM_NAME == "GSW" & Season == 2018)$ShotPct3Pts*3*dplyr::filter(TeamStats, TEAM_NAME == "Hou" & Season == 2018)$PercentageOf3s) + (dplyr::filter(TeamStats, TEAM_NAME == "GSW" & Season == 2018)$ShotPct2Pts*2*dplyr::filter(TeamStats, TEAM_NAME == "Hou" & Season == 2018)$PercentageOf2s),3)` which is an increment, but no by so much.


```{r}
DT::datatable(TeamStats)
```

### Is there any downsides on shooting this many three pointers?

If we look at the proportion of shots a team takes as a proportion of the number of shots their opponent takes as shown in the formula bellow, we see in the following graph (figure 2) that the more threes you take, the less proportions of shots you take, that is, the other team gets more second chances, but the slope of that tendency is lower than the one in Figure 1. That is, usually the proportion of shots you loose by taking more threes is more than offset by the increase in points per shots. 

$$PropShot =\frac{ShotsTaken_{Offense}}{ShotsAllowed_{Defense}}$$

On the other hand it can still change the outcome of a match, if you see in this years [team stats](https://derek-corcoran-barrios.github.io/TeamStats.html), you will see that even when Golden State has the higher Points per shot in the NBA, it only takes `r dplyr::filter(TeamStats, TEAM_NAME == "GSW" & Season == 2018)$PropShot` shots per every opponent shot. That is why we added the column AdjPPS, which is basically the Points per shot you take adjusted by the proportion of shots taken in a match. 

```{r}

prop <- ggplot(TeamStats, aes(x = PropShot, y = PPS)) + geom_point() + geom_smooth(method = "lm") + ggtitle("Figure 2") + geom_text(aes(label=TEAM_NAME),hjust=0, vjust=0) + xlab("Shots taken/shots taken by opponent")

ggplotly(prop)

```
