---
title: "How to use the NetworExtinction Package"
author: "Isidora Avila and Derek Corcoran"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: yes
    fig_caption: yes
bibliography: biblio.bib

---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=6, fig.height=4)
knitr::opts_chunk$set(cache = TRUE, message = FALSE, warning = FALSE)
```

#Introduction

The objective of the NetworkExtinction package is to analyze and visualize the topology of the food-webs. At the same time to analyze and visualize how the food-web responds to the primary removal of species.

The principal index that are incorporated are: 

1. Node numbers: Total number of species in the network [@dunne2002food].

2. Link number: Number of trophic relationships represented in the food web [@dunne2002food].

3. Connectance: Proportion of all possible trophic links that are actually realized [@dunne2002food].

4. Primary removals: Occurs when the researcher intentionally removes a species simulating an extinction.

5. Secondary extinctions: A secondary extinction occurs when a non-basal species loses all of its prey items resulting from the primary removal of another species. In this context, basal species may experience primary removal, but not secondary extinctions [@dunne2002food].

6. Total extinctions: Primary species removal plus secondary extinctions


## How to install the package

```{r, eval=FALSE}
install.packages(NetworkExtinction)
library(NetworkExtinction)
```

This package has four different functions (see above for details):  

* *Mostconnected:* Extinctions from most connected to less connected  

* *ExtinctionOrder:* Extinctions from custom order  

* *RandomExtinctions:* 

* *degree_distribution:* Degree distribution analysis  


## How to create a network object

NetworkExtinction package use a network object to make an analysis.  

For create a network object you can start with a matrix or a edgelist object (for more details see network package [@butts2008network]).

#### A Example
```{r}
#create a matrix
a<- matrix(c(0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,1,0,0,0,1,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,1,1,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), nrow=10, ncol=10)

#the matrix of the data(net) used below is the transpose of a
a<- t(a) #transpose matrix a
a
#create a network
library(network)
net <- as.network(a, loops = TRUE)
net
```


### Extinctions from most connected to less conected

The function to do this analysis is called *Mostconnected*

This function orders the species from the most connected species to the least connected species, using total degree. Then, remove the most connected species in the network and calculate the topological indexes of the network (which were mentioned above) and count how many species have indegree 0 (secondary extinction), without count primary producers. Then, remove the species that were secondarily extinct in the previous step and recalculate which is the new most connected species and so on, until the number of links in the network is zero [@sole2001complexity; @dunne2002food; @dunne2009cascading]

```{r}
library(NetworkExtinction)
data("net")
Mostconnected(Network = net)
```

The result is a dataframe. The first column (Spp) indicates which was the species that was removed as a primary removal. 

The column Secondary_extinctions represent the numbers of the species that are extinguished in a secondary form, and the AccSecondaryExtinction column represents the accumulated secondary extinction.

to plot the results, see function ExtinctionPlot
```{r, fig.cap="Fig 1. The graph shows the number of accumulated secondary extinctions that occur when removing species from the most connected to the least connected"}
data("net")
history <- Mostconnected(Network = net)
ExtinctionPlot(History = history, Variable = "AccSecondaryExtinction")

```


### Extinctions from custom order
The function to do this analysis is called *ExtinctionOrder*

It takes a network extinguishes species using a given order, then it calculates the topological food-web index and the secondary extinctions.

```{r}
data("net")
ExtinctionOrder(Network = net, Order = c(2,4,7))
```

The results it is a dataframe with the topological index of the food-web in each extinction step (like MostConnected results). Additionally, the result generate a plot that show the number of  accumulated secondary extinctions that occur when removing species

### Random extinction
The function to do this analysis is called *RandomExtinctions*
From a network, a certain amount of simulations is done (nsim), in which the species are extinguished in a random order.

the objective is to compare whether the number of secondary extinctions that occur when extinguishing species in a particular order (MostConnected or Extinctionorder) is different from the number of secondary extinctions that occur when executing primary extinction orders in a random manner.

```{r, message=FALSE}
data(net)
RandomExtinctions(Network= net, nsim= 10)
```

The result is a dataframe and a graph. 

*  In the dataframe,  The first column (NumExt) represents the number of extinction that occurs by primary removal. The second one, represents the deviation to the average of the accumulated secondary extinctions that occur when, independently of species identity, a certain number of species is extracted by primary removal. The third one, represents the average of the accumulated secondary extinctions that occur when, regardless of the specie identity, a certain number of species is extracted by primary removal.

*  In the graph, the figure shows the accumulated secondary extinction with their deviation (red shadow) that occur as the number of primary removal increases

###Comparison of Null hypothesis with other extinction histories

By using the *RandomExtincton* function, we create a null hypothesis for us to compare it with either an extinction history generated by the *ExtinctionOrder* function or the *Mostconnected* function. In order to compare the expected extinctions developed by our null hypotehsis with the observed extinction history we developed the *CompareExtinctions* fucntion.

The way to use this function is to first create the extinction history and the null hypothesis and then use the function to compare them:

```{r,message=FALSE, warning=FALSE}
data("net")
History <- ExtinctionOrder(Network = net, Order = c(1,2,3,4,5,6,7,8,9,10))

set.seed(2)
NullHyp <- RandomExtinctions(Network = net, nsim = 1000)

Comparison <- CompareExtinctions(Nullmodel = NullHyp, Hypothesis = History)
```

the resulting object will have a graph with a dashed line showing the observed extincion histoy, and a solid line showing the expected value of secondary extinctions originated at random:

```{r}
Comparison$graph
```

We will also get a Test object which will show the goodness of fit statistics of the comparison:

```{r}
Comparison$Test
```

since the p value is `r round(Comparison$Test$p.value, 2)` which is larger than 0.05, we consider that the generated extinction history is significantly different than the null hypothesis.

### Degree distribution analysis

The function to do this analysis is called *degree_distribution*

degree_distribution function calculate the cumulative distribution of the number of links that each species in the food network has [@estrada2007food]. Then, three expected distribution to the observed distribution, this are:
*  Exponential distribution model 
*  Power-law distribution model  
*  Truncated power-law distribution model  

At the same time, the degree_distribution function calculate the AIC to recognize among the three models mentioned above, which is the one that best fits to the observed distribution.

Finally, the degree_distribution function generate a plot of the observed degree distribution in a log-log scale fitting the three models mentioned above
```{r}
data("net")
degree_distribution(net, name = "Test")
```

The function generate three results:


1.  A list with the observed value of the network degree distribution (column K and Cumulative) and the expected values of the each fitting model (Exp = Exponential model, Power = Power-law model, truncated = Truncated Power-law model). k represent the degree of the network and cumulative the probability that each specie could be have this degree (pk).

2.  A list with the results if the AIC among the three models. In this case the model list are order from the best to the worst model.

3.  A Graph of the observed degree distribution with the fit of the three models.

*   Observation: In the graph, the zero values are not represented but this result are incorporate in the DDvalues result.


Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.


#Bibliography

