---
title: "Spatial Analysis in R"
author: "Derek Corcoran"
date: "`r format(Sys.time(), '%d/%m, %Y')`"
output:
  ioslides_presentation:
    widescreen: true
    incremental: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = FALSE, message = FALSE, cache = FALSE, tidy = TRUE, tidy.opts= list(blank = FALSE, width.cutoff = 60))
library(leaflet)
library(sf)
library(tidyverse)
library(rgdal)
library(raster)
library(rasterVis)
library(rworldxtra)
library(tidyverse)
library(gridExtra)
library(kableExtra)
options("kableExtra.html.bsTable" = T)
```

# What is GIS

## Geographic Information System {.build}

* Spatial representation of data
* Many type of data
    + Remote sensing data (Spectral data, lidar)
    + Modeled datasets (Species distribution models)
    + Expert data (Estimated distribution)
    + Future planning (Systematic conservation planning)
    + Fanatasy maps (Middle-earth, Westeros)
* Two main formats
    + Vectorial data (Shapefiles)
    + Grid data (Rasters)
    
## Vectorial datasets

* Shapefiles (*.shp*)
* **Points** Example `Abundance.csv` (Cities, species presences, events, sampling sites)
* Lines (Rivers, Roads, Paths)
* **Polygons** example `Africa.shp` (Countries, regions, provinces, lakes)
* Packages (sf, sp)

# Polygons

## Vectorial data example

```{r, echo = TRUE}
library(sf)
library(tidyverse)

Africa <- read_sf("Africa.shp")

ggplot() + geom_sf(data = Africa, aes(fill = POP_EST))
```

## Example vectorial data (cont.)

```{r, echo = TRUE, eval = F}
ggplot() + geom_sf(data = Africa, aes(fill = GEO3))
```

# Points example

## Read from csv

```{r, echo = TRUE}
Abundance <- read_csv("Abundance.csv")
Abundance_sf <- Abundance %>% st_as_sf(coords = c(9,10),crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
ggplot() + geom_sf(data = Africa, aes(fill = POP_EST)) + geom_sf(data = Abundance_sf, color = "red")
```

## Plot points using another variable

```{r,echo = T, eval = F}
ggplot() + geom_sf(data = Africa, aes(fill = POP_EST)) + geom_sf(data = Abundance_sf, aes(color = Family))
```

## Plot points using another variable

```{r,echo = T, eval = F}
ggplot() + geom_sf(data = Africa, aes(fill = POP_EST)) + geom_sf(data = Abundance_sf, aes(size = Measurement))
```

# Grid Data

## Grid data

* Similar to a picture
* Each pixel has certain characterstics (resolution, projection, units)
* Example `Clim.tif`


## Read mean temperature and precipitation

```{r, echo = T}
Clim <- stack("Clim.tif")
plot(Clim)
```

## Change names

```{r, echo = T}
names(Clim) <- c("Temp", "Prec")
plot(Clim, colNA = "black")
```

# Summary

## To read

```{r}
DF <- tibble(Action = c("Read"), Shapefile = c("read_sf"), CSV = c("read_csv"), Raster =c("stack/raster"))

kable(DF) %>% kable_styling(bootstrap_options = c("striped", "hover"))
```

# Plot together

## Transform raster to DF

```{r, echo = T}
Clim_DF <- Clim %>% 
  as("SpatialPixelsDataFrame") %>% 
  as.data.frame()

head(Clim_DF)
```

## Plot together

```{r, echo = T}
ggplot() + geom_raster(data = Clim_DF, aes(x = x, y = y, fill = Temp)) + geom_sf(data = Africa, alpha = 0) + geom_sf(data = Abundance_sf, aes(size = Measurement))
```

## Improve scale

```{r, echo = T, eval = F}
ggplot() + geom_raster(data = Clim_DF, aes(x = x, y = y, fill = Temp)) + geom_sf(data = Africa, alpha = 0) + geom_sf(data = Abundance_sf, aes(size = Measurement)) + scale_fill_viridis_c(name = "Temperature")
```

## White background

```{r, echo = T, eval = F}
ggplot() + geom_raster(data = Clim_DF, aes(x = x, y = y, fill = Temp)) + geom_sf(data = Africa, alpha = 0) + geom_sf(data = Abundance_sf, aes(size = Measurement)) + scale_fill_viridis_c(name = "Temperature") +
theme_bw()
```

## Remove axis names

```{r, echo = T}
ggplot() + geom_raster(data = Clim_DF, aes(x = x, y = y, fill = Temp)) + geom_sf(data = Africa, alpha = 0) + geom_sf(data = Abundance_sf, aes(size = Measurement)) + scale_fill_viridis_c(name = "Temperature") +
theme_bw() + labs(y = NULL, x = NULL)
```

# Subsetting and cropping 

## Subsetting Vectorial data

```{r, echo = TRUE}
Southern <- Africa %>% dplyr::filter(GEO3 == "Southern Africa")

ggplot() + geom_sf(data = Southern) + theme_bw()
```

## Subsetting Vectorial data

```{r, echo = TRUE}
SouthCentral <- Africa %>% dplyr::filter(GEO3 %in% c("Southern Africa", "Central Africa"))
```


## Subsetting using numeric variables

```{r}
MedianGDP <- median(Africa$GDP_MD_)

Richest <- Africa %>% dplyr::filter(GDP_MD_ > MedianGDP)
```

# Cropping raster data

## Lets say we just want to see climate from Namibia

* First we create a polygon of Namibia

```{r, echo = T}
Namibia <- Africa %>% dplyr::filter(SOVEREI == "Namibia")
```

* Then crop climate to that

```{r, echo =T}
Clim_Namibia <- Clim %>% crop(Namibia)
```

* If we plot it we can see that it takes all the *Bounding box*

# Modify datasets

## Create vectorial variables with mutate

* Lets generate GDP_Per_Capita
* We will divide `GDP_MD_` by `POP_EST` with mutate

```{r}
Africa <- Africa %>% mutate(GDP_Per_Capita = GDP_MD_/POP_EST)
```

* Now we can see that new variable

```{r}
ggplot() +
  geom_sf(data = Africa, aes(fill = GDP_Per_Capita)) +
  scale_fill_viridis_c() +
  theme_bw()
```

## Create Grid variables

* Temperature in Fahrenheit

```{r, echo = T}
Fahrenheit <- (Clim[[1]]*(9/5)) + 32
```

# Export your data 

## Export your data

* For vectorial data

```{r, echo = T, eval = FALSE}
write_sf(Namibia, "Namibia.shp")
```

* For raster data

```{r, echo = T, eval = FALSE}
writeRaster(Fahrenheit, "Temp_F.tif", format="GTiff", overwrite=TRUE)
```

* **Formats:** .grd, .asc, .sdat, .rst, .nc, .tif, .envi, .bil, .img

# Summary

## To write

```{r}
DF <- tibble(Action = c("Read", "Write"), Shapefile = c("read_sf", "write_sf"), CSV = c("read_csv", "write_sf"), Raster =c("stack/raster", "writeRaster"))

kable(DF) %>% kable_styling(bootstrap_options = c("striped", "hover"))
```

# Resources

## Resources

* [SF](https://r-spatial.github.io/sf/)
* [raster](https://rspatial.org/raster/)
* [Example interactive map using leaflet](https://rpubs.com/derek_corcoran/Mapa_El_Soldado)
* [Interactive maps](https://rstudio.github.io/leaflet/)
* [Spatial Data Science book](https://keen-swartz-3146c4.netlify.app/)
* [Example of webapp in r using shiny](https://www.ecoinformatica.net/GCMcompareR.html)


# If we get the chance

## Model abundance

* Model abundance of Heuglin's white-eye (*Zosterops poliogaster*)

```{r}
ForModel <- Abundance_sf %>% dplyr::filter(Best_guess_binomial == "Zosterops poliogaster")

Area <- Clim %>% crop(ForModel)
```

```{r, echo= F}
Area_DF <- Area %>% as("SpatialPixelsDataFrame") %>% as.data.frame()
Area_Africa <- Africa %>% st_crop(ForModel)
ggplot() + geom_raster(data = Area_DF,aes(x = x, y = y, fill = Temp)) + geom_sf(data =Area_Africa, alpha = 0) + geom_sf(data = ForModel, aes(size = Effort_corrected_measurement)) + theme_bw() + scale_fill_viridis_c() +
  labs(x = NULL, y = NULL)
```

## Prediction

```{r, echo = F}
Area <- Clim %>% crop(ForModel)

Data <- raster::extract(Area, ForModel) %>% as.data.frame()

ForModel <- bind_cols(ForModel, Data)

Model <- glm(Effort_corrected_measurement ~ Temp + Prec, family = poisson, data = ForModel)

Prediction <- predict(Area, Model, type = "response")

Prediction_DF <- Prediction %>% as("SpatialPixelsDataFrame") %>% as.data.frame() %>% rename(Abundance = layer)

ggplot() + geom_raster(data = Prediction_DF,aes(x = x, y = y, fill = Abundance)) + geom_sf(data =Area_Africa, alpha = 0) + geom_sf(data = ForModel, aes(size = Effort_corrected_measurement)) + theme_bw() + scale_fill_viridis_c(trans = 'log') +
  labs(x = NULL, y = NULL)
```

