---
title: "Spatial Analysis in R"
author: "Derek Corcoran"
date: "`r format(Sys.time(), '%d/%m, %Y')`"
output:
  ioslides_presentation:
    widescreen: true
    incremental: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE, tidy = TRUE, tidy.opts= list(blank = FALSE, width.cutoff = 60))
library(leaflet)
library(sf)
library(tidyverse)
library(rgdal)
library(raster)
library(rasterVis)
library(rworldxtra)
library(tidyverse)
library(gridExtra)
library(kableExtra)
options("kableExtra.html.bsTable" = T)
```

# What is GIS

## Geographic Information System {.build}

* Spatial representation of data
* Many type of data
    + Remote sensing data (Spectral data, lidar)
    + Modeled datasets (Species distribution models)
    + Expert data (Estimated distribution)
    + Future planning (Systematic conservation planning)
    + Fanatasy maps (Middle-earth, Westeros)
* Two main formats
    + Vectorial data (Shapefiles)
    + Grid data (Rasters)

## Vectorial datasets

* Shapefiles (*.shp*)
* **Points** Example `Abundance.csv` (Cities, species presences, events, sampling sites)
* Lines (Rivers, Roads, Paths)
* **Polygons** example `Africa.shp` (Countries, regions, provinces, lakes)
* Packages (sf, sp)

# Polygons

## Vectorial data example

```{r, echo = TRUE}
library(sf)
library(tidyverse)

Africa <- read_sf("Africa.shp")

ggplot() + geom_sf(data = Africa, aes(fill = POP_EST))
```

## Example vectorial data (cont.)

```{r, echo = TRUE, eval = F}
ggplot() + geom_sf(data = Africa, aes(fill = GEO3))
```

# Points example

## Read from csv

```{r, echo = TRUE}
Abundance <- read_csv("Abundance.csv")
Abundance_sf <- Abundance %>% st_as_sf(coords = c(9,10),crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
ggplot() + geom_sf(data = Africa, aes(fill = POP_EST)) + geom_sf(data = Abundance_sf, color = "red")
```

## Plot points using another variable

```{r,echo = T, eval = F}
ggplot() + geom_sf(data = Africa, aes(fill = POP_EST)) + geom_sf(data = Abundance_sf, aes(color = Family))
```

## Plot points using another variable

```{r,echo = T, eval = F}
ggplot() + geom_sf(data = Africa, aes(fill = POP_EST)) + geom_sf(data = Abundance_sf, aes(size = Measurement))
```

# Grid Data

## Grid data

* Similar to a picture
* Each pixel has certain characterstics (resolution, projection, units)
* Example `Clim.tif`


## Read mean temperature and precipitation

```{r, echo = T}
Clim <- stack("Clim.tif")
plot(Clim)
```

## Change names

```{r, echo = T}
names(Clim) <- c("Temp", "Prec")
plot(Clim, colNA = "black")
```

# Summary

## To read

```{r}
DF <- tibble(Action = c("Read"), Shapefile = c("read_sf"), CSV = c("read_csv"), Raster =c("stack/raster"))

kable(DF) %>% kable_styling(bootstrap_options = c("striped", "hover"))
```

# Plot together

## Transform raster to DF

```{r, echo = T}
Clim_DF <- Clim %>% 
  as("SpatialPixelsDataFrame") %>% 
  as.data.frame()

head(Clim_DF)
```

## Plot together

```{r, echo = T}
ggplot() + geom_raster(data = Clim_DF, aes(x = x, y = y, fill = Temp)) + geom_sf(data = Africa, alpha = 0) + geom_sf(data = Abundance_sf, aes(size = Measurement))
```

## Improve scale

```{r, echo = T, eval = F}
ggplot() + geom_raster(data = Clim_DF, aes(x = x, y = y, fill = Temp)) + geom_sf(data = Africa, alpha = 0) + geom_sf(data = Abundance_sf, aes(size = Measurement)) + scale_fill_viridis_c(name = "Temperature")
```

## White background

```{r, echo = T, eval = F}
ggplot() + geom_raster(data = Clim_DF, aes(x = x, y = y, fill = Temp)) + geom_sf(data = Africa, alpha = 0) + geom_sf(data = Abundance_sf, aes(size = Measurement)) + scale_fill_viridis_c(name = "Temperature") +
theme_bw()
```

## Remove axis names

```{r, echo = T}
ggplot() + geom_raster(data = Clim_DF, aes(x = x, y = y, fill = Temp)) + geom_sf(data = Africa, alpha = 0) + geom_sf(data = Abundance_sf, aes(size = Measurement)) + scale_fill_viridis_c(name = "Temperature") +
theme_bw() + labs(y = NULL, x = NULL)
```

# Subsetting and cropping 

## Subsetting Vectorial data

```{r, echo = TRUE}
Southern <- Africa %>% dplyr::filter(GEO3 == "Southern Africa")

ggplot() + geom_sf(data = Southern) + theme_bw()
```

## Subsetting Vectorial data

```{r, echo = TRUE}
SouthCentral <- Africa %>% dplyr::filter(GEO3 %in% c("Southern Africa", "Central Africa"))

ggplot() + geom_sf(data = SouthCentral) + theme_bw()
```

## Modificar datos vectoriales

```{r, echo = T}
#Africa <- Africa %>% mutate(Poblacion_mill = POP_EST/1000000)
#ggplot() + geom_sf(data = Africa, aes(fill = Poblacion_mill)) + theme_bw() + scale_fill_viridis_c()
```

## Modificar datos vectoriales (Cont.)


```{r, echo = T}
#Africa <- Africa %>% mutate(PIB_per_Cap = GDP_MD_EST/POP_EST)
#ggplot() + geom_sf(data = Africa, aes(fill = PIB_per_Cap)) + theme_bw()
```

## Exportar y leer datos vectoriales

```{r, echo = T}
#Africa <- Africa %>% dplyr::select(NAME, Poblacion_mill, PIB_per_Cap, GLOCAF)
#write_sf(Africa, "Africa.shp")
#Africa2 <- read_sf("Africa.shp")
#ggplot() + geom_sf(data = Africa2, aes(fill = GLOCAF)) + #theme_dark() 
```


## Algunas bases de datos vectoriales

* [Datos vectoriales del mundo](http://www.diva-gis.org/gdata)
* [Datos vectoriales de Chile](https://www.bcn.cl/siit/mapas_vectoriales)


## Algunas bases de datos vectoriales

* En raster función `getData("GADM")`

```{r, echo = T}
#Peru <- getData(name = "GADM", country = "PER", level = 1) %>% #st_as_sf()
#ggplot() + geom_sf(data = Peru) + theme(legend.position = "none")
```


# Raster


```{r, echo = T}
#Prec <- getData("worldclim", res = 10, var = "prec") 
#Prec
```