---
title: "Leaflet miniworkshop"
author: "Derek Corcoran"
date: "`r format(Sys.time(), '%d/%m, %Y')`"
output:
  ioslides_presentation:
    widescreen: true
    incremental: true
---

```{r setup, include=FALSE}
#remotes::install_github("rlesur/klippy")

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = FALSE, tidy = TRUE, tidy.opts= list(blank = FALSE, width.cutoff = 60))
library(leaflet)
library(kableExtra)

options("kableExtra.html.bsTable" = T)
```



## load packages

```{r, echo = TRUE, eval= FALSE}
library(leaflet)
```

## We will start by using some points

```{r}
Bombus <- read.csv("https://raw.githubusercontent.com/derek-corcoran-barrios/derek-corcoran-barrios.github.io/master/Presentaciones_Espacial/Species.csv")
Bombus <- Bombus[,c(3,5:7,9)]
colnames(Bombus)[5] <- "Abundance"
```

```{r, echo = FALSE}
 knitr::kable(Bombus) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "350px")
```


## Lets just add points to a map

```{r}
M <- leaflet() %>% 
  addTiles() %>% 
  addCircles(data = Bombus, lng = ~Longitude, ~Latitude) 
```

* Otras opciones
    + addAwesomeMarkers
    + addCircleMarkers
    + addMarkers

## We can add some controls

```{r}
M %>% addMeasure(primaryLengthUnit = "meters", primaryAreaUnit = "hectare")
```


## Change color by species

* Generate a pallete

```{r}
pal <- colorFactor(c("red", "blue"), domain = unique(Bombus$species))
```

* Add it to the species with color and fillColor

```{r}
M <- leaflet() %>% 
    addTiles() %>% 
    addCircles(data = Bombus, lng = ~Longitude, ~Latitude, color = ~pal(species), fillColor = ~pal(species))
```

* Add Legend

```{r}
M <- M %>% addLegend(pal = pal, values = unique(Bombus$species))
```


## Change color

```{r, echo = FALSE}
M
```

## Add popups and labels

```{r}
M <- leaflet() %>% 
    addTiles() %>% 
    addCircles(data = Bombus, lng = ~Longitude, ~Latitude, 
               color = ~pal(species), fillColor = ~pal(species),
               popup = ~Site_name, label = ~Abundance)
```


# Problem Species overlap

## We can change size for abundace

* Switched to CircleMarkers

```{r}
  M <- leaflet() %>% 
    addTiles() %>% 
    addCircleMarkers(data = Bombus, lng = ~Longitude, ~Latitude, 
                     color = ~pal(species), fillColor = ~pal(species),
                     popup = ~Site_name, label = ~Abundance,
                     radius = 10*(Bombus$Abundance/max(Bombus$Abundance)))
  
  M %>% addLegend(pal = pal, values = unique(Bombus$species))
```

